---
title: "Presidential Forecasting"
author: "Ike"
date: "4/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

INTERESTING FINDINGS
-HRC recieved 22x as many votes as DJT in DC
-only 6 states becoming more moderate from linear extrapolation

-machine learning model to forecast election results at the state level; include confidence interval
-can be used, for example, to find probability that PA goes blue in 2028--will obviously be less accurate over time
-data taken from MIT Election Data and Science Lab, 2017, "U.S. President 1976-2016", https://doi.org/10.7910/DVN/42MVDX, Harvard Dataverse, V5, UNF:6:Mw0hOUHAijKPTVRAe5jJvg== [fileUNF]

```{r read_file}
options(digits = 3)
library(tidyverse)
filename <- '1976-2016-president.csv'
read_lines(filename, n_max = 3)
file <- read.csv(filename)
```

-filter to only rep and dem for each state
-very small inaccuracy introduced by throwing away write-ins--even if for major party candidate

```{r simlify_data}
# filter to only rep and dem for each state; very small inaccuracy introduced by throwing away write-ins--even if for major party candidate
# wild there isn't a better not in function
'%ni%' <- Negate('%in%')
dem <- file %>% filter(party %in% c('democrat', "democratic-farmer-labor") & candidate %ni% c('Other', '') & writein == FALSE) %>% select(year, state_po, candidatevotes, totalvotes) %>% mutate(party = 'democrat')
rep <- file %>% filter(party == 'republican' & writein == FALSE) %>% select(year, state_po, party, candidatevotes, totalvotes)
votes_simple <- data.frame(year = dem$year, state = dem$state_po, percent_dem = dem$candidatevotes / (rep$candidatevotes + dem$candidatevotes))
```

-plot of every state over time
-very busy and I don't know if it gives that much information
-larger range in data which hints at increased polarization. can't really see divergence on a state-by-state basis but again, a very busy plot

```{r plot_all_states, include = FALSE}
ggplot(votes_simple, aes(year, percent_dem, col = state)) + geom_line() +
  geom_text(aes(label = state), size = 3.5) +
  theme(legend.position = "none")
```

-linear model fit for each state
-interesting that GA and TX are both becoming more conservative even with more liberal trends in the last several elections due to 1976 and 1980 being liberal years. could begin in 1984 but that feels like cherry picking. including an x^2 term will most likely remove this.
-should also look into something that gives an increased weight to more recent data
-note: includes confidence interval not prediction interval. looks at averages, not individual predictions

```{r plot_linear}
setwd('C:/Users/ikeur/OneDrive/Documents/learn r/presidential_forecasting/graphs/states_linear')
sts <- votes_simple %>% filter(year == 1976) %>% .$state
for (s in sts){
  name <- paste(s, 'linear.jpg', sep = '')
  jpeg(filename = name)
  filt <- votes_simple %>% filter(state == s)
  lin <- lm(percent_dem ~ year, data = filt)
  new.dat <- data.frame(year = c(1976, 2016))
  pre <- predict(lin, newdata = new.dat, interval = 'confidence')
  include_50 <- (max(pre, filt$percent_dem) > .48 & min(pre, filt$percent_dem) < .52)
  st <- filt %>% ggplot(aes(year, percent_dem)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle(s) + 
    theme(plot.title = element_text(hjust = 0.5))
  if (include_50){
    plot(st + geom_segment(aes(x = 1976, y = .5, xend = 2016, yend = .5), linetype='dashed'))
  } else {
    plot(st)
  }
  dev.off()
}
```